---
title: "SuperMarketSales"
author: "Aditya Pansari"
date: "2024-01-26"
output: html_document
---
# Super Market Sales
### Let's see what are the pre installed libraries in the system
```{r}
installed.packages()

```
### Let's see what are the packages priorly loaded
```{r}
search()
```
### Let's set the file path and load the data into the variable
```{r}
file_path=("D:/SuperMarketSales/supermarket_sales - Sheet1.csv")
data<-read.csv(file_path) 
```
### let's get the summary of the data
```{r}
summary(data)
```
###Let's look into the fisrt 6 Rows
```{r}
head(data)
```
### viewing the class of each variables
```{r}
class(data$Invoice.ID)
class(data$Branch)
class(data$City)
class(data$Customer.type)
class(data$Gender)
class(data$Product.line)
class(data$Unit.price)
class(data$Quantity)
class(data$Tax.5.)
class(data$Total)
class(data$Date)
class(data$Time)
class(data$Payment)
class(data$cogs)
class(data$gross.margin.percentage)
class(data$gross.income)
class(data$Rating)
```
```{r}
class(data)
```
```{r}
is.blank <- function(x) {
  ifelse(is.na(x), TRUE, ifelse(is.character(x), trimws(x) == "", FALSE))}
```
### Replace blank values with NA in the entire data frame
```{r}
data[data == ""] <- NA
data[sapply(data, is.blank)] <- NA
```

```{r}
str(data)
```

```{r}
duplicates <- data[duplicated(data), ]
print(duplicates)
print(nrow(duplicates))
```
#Checking for NAs within each column/variable
```{r}
sum_na <- colSums(is.na(data))
```
# Display the results
```{r}
print(sum_na)
```
```{r}
library(tidyr)      #Data manipulation and visualization
library(dplyr)      # Data manipulation
library(ggplot2)    # Data visualization
library(lubridate)  # Date and time manipulation
library(readxl)
```

### Bar plot for the distribution of customer types
```{r}
ggplot(data, aes(x = Customer.type, fill = Customer.type)) +
  geom_bar() +
  geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5, color = 'black', size = 4) +  
  ggtitle("Distribution of Customer Types")

```

###  Barplot of gender distribution
```{r}
ggplot(data, aes(x = Gender, fill = Gender)) +
  geom_bar() +
  geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5, color = 'black', size = 4) +  
  ggtitle("Gender Distribution of Customers")
```

### Product Line categories distribution barplot
```{r}
ggplot(data, aes(x = Product.line, fill = Product.line)) +
  geom_bar() +
  geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5, color = 'black', size = 4) +  
  ggtitle("Product Line Distribution") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # rotates the labels on the x axis
```

### Payment methods used by customers
```{r}
ggplot(data, aes(x = Payment, fill = Payment)) +
  geom_bar() +
  geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5, color = 'black', size = 4) +  
  ggtitle("Payment Method Distribution")
```

### Ratings of customer shopping experiences
```{r}
ggplot(data, aes(x = Rating, fill = factor(Rating))) +
  geom_bar(width = 0.07) + 
  ggtitle("Distribution of Ratings")
```

```{r}
data %>% filter(Tax.5. == 0.5085)
```

```{r}
data %>%
group_by(Product.line) %>%
filter(Tax.5. >= 25.00)
```

```{r}
Gender_category <- data %>%
group_by(Gender, Product.line) %>%
summarise(count= n())
Gender_category
```

```{r}
library(ggrepel)
```

```{r}
ggplot(data = Gender_category, aes(x = Product.line, y= count)) + geom_bar(stat = "identity", width = 0.6, aes(fill = Gender)) + coord_flip() +
 geom_text_repel(aes(label = count)) + theme_minimal()+ theme(axis.text.x = element_text(angle = 90)) + scale_fill_brewer(palette="Spectral")
```

```{r}
result <- data %>%
group_by(Product.line ) %>%
summarise("Cost_of_goods_sold" = mean(cogs))
result
```

```{r}
branchcosts <- data %>%
group_by(Product.line,Branch ) %>%
summarise("Cost_of_goods_sold" = mean(cogs))
branchcosts
```

```{r}
ggplot(data=result, aes( x= Product.line, y= Cost_of_goods_sold )) +geom_point(aes(colour = Product.line)) + theme(axis.text.x = element_text(angle = 60)) + scale_fill_brewer(palette="Spectral")
```

```{r}
ggplot(data=result, aes( x= Product.line,y=Cost_of_goods_sold)) +geom_bar(stat = "identity", width = 0.6,aes(fill = Product.line)) +
labs(x = "Categories", y = "COGS", title = "Cost per category") +theme_minimal() + theme(axis.text.x = element_text(angle = 90)) + scale_fill_brewer(palette="Spectral")
```

```{r}
branch_products <- data %>%
group_by(Branch, Product.line) %>%
summarise(count= n())
branch_products
```

```{r}
data %>%
group_by(Branch) %>%
summarise("Quantity of goods sold per branch" = sum(Quantity))
```

```{r}
ggplot(data = branch_products, aes(x= Branch, y= count, fill = Product.line)) + geom_bar(stat = "identity") +
  labs(title = "Branch wise sale of categories",
       x = "Branch",
       y = "Units sold",
       fill = "Legend")   + theme(axis.text.x = element_text(angle = 90)) +theme_minimal() + scale_fill_brewer(palette="Spectral")
```

```{r}
data %>%
group_by(Customer.type) %>%
summarise("Purchasequantity" = sum(Total))
```

```{r}
data %>%
group_by(Customer.type) %>%
summarise(count = n())
```

```{r}
Citymember <- data %>%
group_by(City,Customer.type) %>%
summarise(count=n())
Citymember
```

```{r}
ggplot(data = Citymember, aes(x= City, y= count, fill = Customer.type)) + geom_bar(stat = "identity") +
  labs(title = "Member/customer by cities",
       x = "City",
       y = "Customers",
       fill = "Legend") + geom_text_repel(aes(label = count)) + theme_minimal() + scale_fill_brewer(palette="Spectral")
```

```{r}
Totalrevenue <- data %>%
group_by(City) %>%
summarise("Gross_revenue" = sum(Total))
Totalrevenue
```

```{r}
ggplot(data= Totalrevenue, aes(x= City, y= Gross_revenue))+ geom_bar(stat= "identity", width = 0.5, fill = "#75ae97", colour ="White")
```


```{r}
COGScity <- data%>%
group_by(City) %>%
summarise("Cost_of_goods_sold" = sum(cogs))
COGScity
```

```{r}
citinm <- unique(data$City)
citinm
```

```{r}
citystats <- data.frame(City_name = citinm,Gross_revenue =Totalrevenue$Gross_revenue,Cost_of_good_sold =COGScity$Cost_of_goods_sold)
citystats
```

```{r}
citystats$grossprofit <- citystats$Gross_revenue - citystats$Cost_of_good_sold
citystats
```


```{r}
ggplot(data = citystats, aes(x= City_name, y= grossprofit)) +geom_bar(stat = "identity", width = 0.5, fill = "#f56400") +theme_minimal()
```

```{r}
ggplot(data= data, aes(x= Total)) + geom_histogram(binwidth = 15, fill= "#5B6A87", colour = "Black") +
xlab("Total bill prices") + ylab("Count of bills") + theme(axis.title.x = element_text(colour = "red", size = 10), axis.title.y = element_text(colour = "Red", size = 10)) + ggtitle("Total bill")
```

```{r}

```

